A Zero-Knowledge Protocol for Keystroke Authentication

### 2014-12-15 paper/paper.tex
draft of paper

Many common websites such as Facebook and Dropbox are already encouraging the use two-factor authentication to subvert these issues. These forms of authentication usually involve confirming the possession of a piece of hardware. Instead, we suggest an alternate authentication factor that does not require possession of additional hardware. Our authentication scheme is based on keystroke dynamics; the user is authenticated based on the unique timing of his/her keystrokes when typing a particular phrase (such as their name). When typing, each individual has their own typing pattern consisting of lag and dwell time (how long a key is pressed) and rhythms (typing certain clusters of letters faster).  This typing pattern translates into a typing signature which can be measured and used as a form of biometric data. Instead of sending this naked data through vulnerable communication channels, we design a zero-knowledge protocol where the server verify the identity of a user without ever having to exchange the biometric data itself. In this way, if there were ever a breach of security, such as the case with Target or J.P. Morgan Chase, no useful information is leaked.
\end{abstract}

@@

We propose an authentication protocol based on keystroke dynamics, the unique timings associated with the way that a user types.  Because each individual prefers their own hand layout on the keyboard, their finger movements will be correspondingly different.  For example, if Albert keeps his left pinky on the \textit{shift} key in rest position, then the \textit{a} key will be covered by his ring finger and his index finger will be split between \textit{d} and \textit{f}.  This will affect his typing speed of certain words, as opposed to Bob who uses his left pinky for the \textit{a} key so has both index and middle fingers available for \textit{d} and \textit{f}.  In this paper, we exploit this fact and use it as the authentication factor, however this protocol could easily be extended to incorporate more factors. As stated before, keystroke dynamics act as a hardware factor without requiring users to carry any piece of physical hardware with them other than their hands. Login on an untrusted machine is therefore equally as secure as login from a trusted one, but it also requires no extra steps on behalf of the user.

@@

\subsection{Zero-Knowledge Proofs for Authentication}
\label{fig:bg-zkp}
In the classical definition of a zero-knowledge proof, a prover $P$ must convince a less powerful verifier $V$ of some fact without giving away any information that would allow $V$ to prove it himself~\cite{GMW86}. An intuitive example of this is the following:

@@

\subsection{Fuzzy Extractors}
\label{sec:fuzzy}
Fuzzy extractors are used to extract a uniformly random string $S$ from a noisy input $\mathfrak{B}$, also known as a biometric template. A similar input $\tilde{\mathfrak{B}}$ will product the exact same string $S$~\cite{conf/csreaSAM/AlvarezEA09}. In the case of keystroke based authentication, $S$ is a user's unique signature which can be extracted from a noisy input. It is then possible to prove knowledge of the secret value $S$ using a zero-knowledge proof. The following construction for a fuzzy extractor is due to \'Alvarez, Encinas, Hern\'andez, and Sanchez-Avila.

@@

We wish to examine the feasibility of keystroke based authentication from both a theoretical and empirical perspective. Two protocols are presented. The first procotol is extremely insecure, but acts as a proof of concept and comparison baseline in terms of accuracy and security. This protocol has been implemented as a simple web app. The second protocol is a zero-knowledge version of the first one that relies on an extractor to obtain a secret from a noisy input. The secret is never stored or communicated across a network; instead knowledge of the secret is proven using a zero-knowledge protocol. This protocol has been implemented using a rounding extractor and the feasibility of using the fuzzy extractor that is described in Section~\ref{sec:fuzzy} is also examined.

When discussing these protocols, the words \textit{prover} and \textit{client} are used almost interchangeably as are the words \textit{verifier} and \textit{server}. The client is an implementation of a prover and the server is an implementation of a verifier. The client is naturally more `powerful' than the server as the client knows the secret keystroke timings (or can at least extract them from a noisy input). Both the regular and the zero-knowledge versions of the web app have two capabilites: account creation and login. The account creation phase requires a user to type their name in a text box. The keystroke timings are recorded and communicated to the server. The server then stores some data about the keystrokes in a database. When a user tries to log in, he also simply types his name in a text box. The name that he typed is looked up in the database and then his keystroke timings are compared against the ones that are stored. Note that the word \emph{compare} is used loosely here; in the zero-knowledge protocol, no direct comparison occurs per se.  Instead we measure the difference between the two signautres, allowing some error tolerance.   The user experience is identical in the regular and zero-knowledge versions, although the implementation is quite different.

\subsection{The Basic Protocol}

The basic protocol authenticates users by directly comparing the given keystroke vector to the expected keystroke vector. This protocol is not designed with security in mind; any adversary that can see the information exchange between the prover and the verifier will be able to directly discern the keystroke vectors. More concretely, anyone evesdropping on the communications over the netork can see these values communicated in the clear. This protocol is not meant to be used in a real implementation of the system, but instead it is meant to be a point of comparison for the zero knowledge version.

@@

Let $V=v_1v_2\ldots v_n$ be the real keystroke vector for user $A$ and $\tilde V = \tilde v_1\tilde v_2\ldots\tilde v_n$ be the keystroke vector associated with a login attempt by someone claiming to be $A$. Account creation in the basic protocol is very simple. First, the prover sends a pair $(A,V)$ containing the user's name and keystroke timings to the verifier. The verifier then stores this information in its knowledge bank. This protocol is outlined in Figure~\ref{fig:basic_new}. Authentication (as seen in Figure~\ref{fig:basic_auth}) is then performed by applying a comparison function $c$ to the expected keystroke vector for user $A$ and the login vector. The comparison function that is used in the actual implementation is defined as follows:
$$\begin{array}{ccc}
c(V,\tilde V) = \begin{cases}
  1, & d(V, \tilde V) \le \epsilon \\
  0, & \text{otherwise}
\end{cases}
& \;\;\;\;\; &
d(V,\tilde V) = \sqrt{\frac1n\sum_{i=1}^n(v_i - \tilde v_i)^2}
\end{array}$$
Above, $\epsilon$ is a constant error tolerance which is carefully chosen considering the tradeoffs between security and ease of login. This tradeoff will be examined closely in section~\ref{sec:results}.

@@

\subsection{The Zero-Knowledge Protocol}
The zero knowledge builds on top of the basic protocol with the added constraint that no meaningful information about the keystroke vector is ever transferred accross the network. This also means that the representation of the keystroke vector that is stored in the database does not leak any information about the keystroke vector itself. If the database were completely compromised, there would thus be no breach in security; the information in the database can be used to verify a user's identity and nothing more.

\begin{figure}
\centering
\fbox{
\begin{tabular}{m{2.5in}m{.5in}m{2.5in}}
\textbf{Prover} && \textbf{Verifier}\\
Choose some generator $g$ & $\xrightarrow{g}$ & Remember $g$\\
User types name ($A$) in web client. Keystroke vector ($V$) is recorded. Let $h = g^{f(V)}$, where $f(V)$ is the fuzzy extraction of $V$.   & $\xrightarrow{h}$ & $(A,g,h)$ is stored in the database
\end{tabular}}
\caption{Account creation in the zero-knowledge protocol}
\label{fig:zk_new}
\end{figure}

\begin{figure}
\centering
\fbox{
\begin{tabular}{m{2.5in}m{.5in}m{2.5in}}
\textbf{Prover} && \textbf{Verifier}\\
User types name ($A$) in web client. Keystroke vector ($\tilde V$) is recorded &$\stackrel{A}{\longrightarrow}$ & \\
&$\stackrel{g}\longleftarrow$ & Lookup key $A$ in the database to get $g$ and $h$\\
Choose $r\in_R\mathbb{Z}_g$, let $a = g^r$ & $\stackrel{a}\longrightarrow$ & \\
& $\stackrel{b}\longleftarrow$ & Choose $b\in_R\mathbb{Z}_g$ \\
Let $c = r + f(\tilde V)\cdot b$ & $\stackrel{c}\longrightarrow$ & \\
Report decision & $\stackrel{ah^b \stackrel{?}{=} g^c}\longleftarrow$ &
\end{tabular}}
\caption{Authentication in the zero-knowledge protocol}
\label{fig:zk_auth}
\end{figure}
\newpage
\section{Results}
\label{sec:results}
We evaluate performance based on two main metrics: precision (how easily a person can repeat their own typing pattern), and protection (how easily an imposter can forget the typing pattern). As stated in Section~\ref{sec:bg-keystroke}, this adversary may not be an actual person; it may be a bot that is programmed to elicit the keystroke timing.  We test the performance of our application with four participants, all with varying typing speeds, preferred hand layouts, and dominant hands.  We measured $\epsilon^{'}$ which is the weighted difference in average typing speed the target user was typing at, compared to our threshold level of $\epsilon=40$. We then measured how consistently the target user could replicate that result, and measured that as a percentage. We then allowed the three other participants to attempt to forge the typing signature, having not seen the target user's test trials.

\begin{table}[h!]
\centering
\begin{tabular}{||c c c c||} 
 \hline
 Length of Name & $\epsilon^{'}$ & Success by User (\%) & Success by Imposter (\%) \\ [0.5ex] 
 \hline\hline
 3 & 112 & 4 & 0 \\ 
 5 & 59 & 10 & 5 \\
 8 & 47 & 85 & 23 \\ 
 11 & 34 & 87 & 43 \\
 13 & 32 & 87 & 35 \\ [1ex] 
 \hline
\end{tabular}
\caption{Keystroke variation by length of name.}
\label{table:1}
\end{table}

Table \ref{table:1} documents repeated test trials with $\delta=0.05$.  Trials were conducted between 20 and 40 times with varying word lengths.  We can see the trend of successful authentications with increasing word length.  This is most likely attributed to the fact that typing patterns are more distinct with longer pieces of text.  On the flip side, imposters are also more likely to successfully impersonate the target user with a longer text, since our algorithm does not differentiate where the difference in time accumulated; in other words, if an imposter's speed did not match in the beginning, (s)he can make up for it by matching in the end.  The error would be spread across the length of the text and therefore validate an imposter.

However, with shorter words, our implementation is very unforgiving.  Since there are fewer keystrokes to average out the mistakes, timing is extremely sensitive.  In practice, however, this is less of a concern, since virtually no one has a combined forename and surname of less than three characters.   As the words become longer, there is more tolerance for fluctuations.

There exists a non-insignificant difference between the success rates of target users and their impersonators.  This affirms the fact that keystroke dynamics are variable enough and have sufficient entropy to be used as authentication tools.  Barring anomalies with short phrases, the difference in patterns become apparent, and also the increase between a phrase of 11 characters and one of 13.

@@

\section{Future Work}
\label{sec:future}
It is evident that the implementation, though viable as a proof of concept, is nowhere near robust enough to be used in practice.  Currently, for longer names, the prototype has bounded probability of success; it correctly verifies the correct user a significant portion of time (barring human induced errors such as typing anomalies), but also verifies an imposter over $40\%$ of the time.  This can be amended by adjusting $\delta$, the fault tolerance per keystroke.  The idea is to find the optimal value of $\delta$ and $\epsilon$ such that the real users can be validated but an imposted can not.  We also would need to adjust our algorithm for those with shorter names.  As can be seen in the previous section, it is extremely difficult for a three letter name to produce a match regardless of the user.  Moving forward, we would need to account for these edge cases, and perhaps impose a character minimum, similar to passwords today.

It is also important to note that this protocol is very keyboard dependent. This means that this method of verification would be impossible on say mobile devices, since people generally use their thumbs to type on smaller keyboards.  Also, a signature on a QWERTY keyboard would be significantly different than one on a French AZERTY keyboard or a Dvorak. In practice, this is less of a concern, since most people form typing habits on one particular keyboard layout, and the variety of keyboard layouts only provide an extra layer of security.

We would also need to test the protocol on many more words of varying lexicographical distances. Our demonstration used only five words, which is subject to extreme bias, since characters in the words chosen may have been clustered in a way that changed the difficulty of typing that one particular word.  The currently observed trends may differ when a larger sample size of words and participants is used.

Though the implementation is not yet consistent, it provides a solid proof of concept for the potential of keystroke authentication.  As we improve the algorithm to weigh more characteristics of a person's typing pattern (such as measuring the actuation force, duration), secure authentication for both parties will be possible. Because of the nature of our zero-knowledge scheme, the users also do not have to worry about any security breaches in the communication channels or in the database. Just like a written signature, the typing signature will be unique and secure.

>>>>

Many common websites such as Facebook and Dropbox are already encouraging the use two-factor authentication to subvert these issues. These forms of authentication usually involve confirming the possession of a piece of hardware. Instead, we suggest an alternate authentication factor that does not require possession of additional hardware. Our authentication scheme is based on keystroke dynamics; the user is authenticated based on the unique timing of his keystrokes when typing a particular phrase (such as his name). When typing, each individual has his own typing pattern consisting of lag and dwell time (how long a key is pressed) and rhythms (typing certain clusters of letters faster).  This typing pattern translates into a typing signature which can be measured and used as a form of biometric data. Instead of sending this naked data through vulnerable communication channels, we design a zero-knowledge protocol where the server verify the identity of a user without ever having to exchange the biometric data itself. In this way, if there were ever a breach of security, such as the case with Target or J.P. Morgan Chase, no useful information is leaked.
\end{abstract}

@@

We propose an authentication protocol based on keystroke dynamics, the unique timings associated with the way that a user types.  Because each individual prefers his own hand layout on the keyboard, his finger movements will be distinct.  For example, if Albert keeps his left pinky on the \textit{shift} key in rest position, then the \textit{a} key will be covered by his ring finger and his index finger will be split between \textit{d} and \textit{f}.  This will affect his typing speed of certain words, as opposed to Bob who uses his left pinky for the \textit{a} and thus has both his index and middle fingers available for \textit{d} and \textit{f}.  In this paper, we exploit this fact and use it as the sole authentication factor, however this protocol could easily be extended to incorporate more factors. As stated before, keystroke dynamics act as a hardware factor without requiring users to carry any piece of physical hardware with them other than their hands. Login on an untrusted machine is therefore equally as secure as login from a trusted one, but it also requires no extra steps on behalf of the user.

@@

\subsection{Zero-Knowledge Proofs for Authentication}
\label{sec:bg-zkp}
In the classical definition of a zero-knowledge proof, a prover $P$ must convince a less powerful verifier $V$ of some fact without giving away any information that would allow $V$ to prove it himself~\cite{GMW86}. An intuitive example of this is the following:

@@

\subsection{Fuzzy Extractors}
\label{sec:bg-fuzzy}
Fuzzy extractors are used to extract a uniformly random string $S$ from a noisy input $\mathfrak{B}$, also known as a biometric template. A similar input $\tilde{\mathfrak{B}}$ will product the exact same string $S$~\cite{conf/csreaSAM/AlvarezEA09}. In the case of keystroke based authentication, $S$ is a user's unique signature which can be extracted from a noisy input. It is then possible to prove knowledge of the secret value $S$ using a zero-knowledge proof. The following construction for a fuzzy extractor is due to \'Alvarez, Encinas, Hern\'andez, and Sanchez-Avila.

@@

We wish to examine the feasibility of keystroke based authentication from both a theoretical and empirical perspective. Two protocols are presented. The first procotol is extremely insecure, but acts as a proof of concept and comparison baseline in terms of accuracy and security. This protocol has been implemented as a simple web app. The second protocol is a zero-knowledge version of the first one that relies on an extractor to obtain a secret from a noisy input. The secret is never stored or communicated across a network; instead knowledge of the secret is proven using a zero-knowledge protocol. This protocol has been implemented using a rounding extractor and the feasibility of using the fuzzy extractor that is described in Section~\ref{sec:bg-fuzzy} is also examined.

When discussing these protocols, the words \textit{prover} and \textit{client} are used almost interchangeably as are the words \textit{verifier} and \textit{server}. The client is an implementation of a prover and the server is an implementation of a verifier. The client is naturally more `powerful' than the server as the client knows the secret keystroke timings (or can at least extract them from a noisy input). Both the regular and the zero-knowledge versions of the web app have two capabilites: account creation and login. The account creation phase requires a user to type his name in a text box. The keystroke timings are recorded and communicated to the server. The server then stores some data about the keystrokes in a database. When a user tries to log in, he also simply types his name in a text box. The name that he typed is looked up in the database and then his keystroke timings are compared against the ones that are stored. Note that the word \emph{compare} is used loosely here; in the zero-knowledge protocol, no direct comparison occurs per se.  Instead, the difference between the two signatures is measured by some specialized technique. The user experience is identical (aside from login accuracy) in the regular and zero-knowledge versions, although the implementation is quite different.

\subsection{The Basic Protocol}
\label{sec:basic-prot}
The basic protocol authenticates users by directly comparing the given keystroke vector to the expected keystroke vector. This protocol is not designed with security in mind; any adversary that can see the information exchange between the prover and the verifier will be able to directly discern the keystroke vectors. More concretely, anyone evesdropping on the communications over the network can see these values communicated in the clear. This protocol is not meant to be used in a real implementation of the system, but instead it is meant to be a point of comparison for the zero knowledge version.

@@

Let $V=v_1v_2\ldots v_n$ be the real keystroke vector for user $A$ and $\tilde V = \tilde v_1\tilde v_2\ldots\tilde v_n$ be the keystroke vector associated with a login attempt by someone claiming to be $A$. Account creation in the basic protocol is very simple. First, the prover sends a pair $(A,V)$ containing the user's name and keystroke timings to the verifier. The verifier then stores this information in its knowledge bank. This protocol is outlined in Figure~\ref{fig:basic_new}. Authentication (as seen in Figure~\ref{fig:basic_auth}) is then performed by applying a comparison function $c$ to the expected keystroke vector for user $A$ and the login vector. The comparison function that is used in the actual implementation is defined as follows:
$$\begin{array}{ccc}
c(V,\tilde V) = \begin{cases}
  1, & d(V, \tilde V) \le \epsilon \\
  0, & \text{otherwise}
\end{cases}
& \;\;\;\;\; &
d(V,\tilde V) = \sqrt{\frac1n\sum_{i=1}^n(v_i - \tilde v_i)^2}
\end{array}$$
Above, $\epsilon$ is a constant error tolerance which is carefully chosen considering the tradeoffs between security and ease of login. This tradeoff will be examined in Section~\ref{sec:results}.

@@

\subsection{The Zero-Knowledge Protocol}
\label{sec:zkp-prot}
The zero knowledge builds on top of the basic protocol with the added constraint that no meaningful information about the keystroke vector is ever transferred accross the network. This also means that the representation of the keystroke vector that is stored in the database does not leak any information about the keystroke vector itself. If the database were completely compromised, there would thus be no breach in security; the information in the database can be used to verify a user's identity and nothing more.

The login process is detailed in Figure~\ref{fig:zk_new}. In this figure, $f$ is the extractor being used. The protocol itself is abstract in that $f$ could represent any extractor; for example, the fuzzy extractor described in Section~\ref{sec:bg-fuzzy}. The protocol is based on the zero-knowledge proof presented in Section~\ref{sec:bg-zkp}. The client first chooses a generator $g$ and a prime $p$, and then sends the value $g^{f(V)}\mod p$ to the server as the user's comparison point. Recall from Section~\ref{sec:bg-zkp} that it is safe to transmit this value because of the assumption that taking the discrete log is hard.

\begin{figure}[h!]
\centering
\fbox{
\begin{tabular}{m{2.5in}m{.5in}m{2.5in}}
\textbf{Prover} && \textbf{Verifier}\\
Choose some generator $g$ & $\xrightarrow{g}$ & Remember $g$\\
User types name ($A$) in web client. Keystroke vector ($V$) is recorded. Let $h = g^{f(V)}$, where $f(V)$ is the fuzzy extraction of $V$.   & $\xrightarrow{h}$ & $(A,g,h)$ is stored in the database
\end{tabular}}
\caption{Account creation in the zero-knowledge protocol}
\label{fig:zk_new}
\end{figure}

The authentication protocol is shown in Figure~\ref{fig:zk_auth}. Since the client in this implementation is a web page that may be accessed from any computer, information about a particular user cannot be stored and thus the value of $g$ must be provided by the server after discovering the user's name. Aside from this caveat, the procotol is identical to the zero knowledge proof presented in Section~\ref{sec:bg-zkp}.

\begin{figure}[h!]
\centering
\fbox{
\begin{tabular}{m{2.5in}m{.5in}m{2.5in}}
\textbf{Prover} && \textbf{Verifier}\\
User types name ($A$) in web client. Keystroke vector ($\tilde V$) is recorded &$\stackrel{A}{\longrightarrow}$ & \\
&$\stackrel{g}\longleftarrow$ & Lookup key $A$ in the database to get $g$ and $h$\\
Choose $r\in_R\mathbb{Z}_g$, let $a = g^r$ & $\stackrel{a}\longrightarrow$ & \\
& $\stackrel{b}\longleftarrow$ & Choose $b\in_R\mathbb{Z}_g$ \\
Let $c = r + f(\tilde V)\cdot b$ & $\stackrel{c}\longrightarrow$ & \\
Report decision & $\stackrel{ah^b \stackrel{?}{=} g^c}\longleftarrow$ &
\end{tabular}}
\caption{Authentication in the zero-knowledge protocol}
\label{fig:zk_auth}
\end{figure}

In the implementation, each keystroke is hashed in a locally sensitive way. First, the keystroke timing is rounded within some small error tolerance and then a hash function is applied to the rounded value. The purpose of the hash function is to scatter the data points in the case that a poor generator is used. If $g^x$ is less than $p$, then it is easy to recover $x$ by taking the base $g$ log of the exponentiated value. A rounding parameter $0 < \delta \le 1$ must be chosen such that it is relatively easy for a person to produce each keystroke while preserving a $\delta$ fraction of the information stored in each keystroke timing, but it is hard for anyone else to achieve this. The extractor function $f$ is defined as follows:

$$f(V) = \{g(v_1),g(v_2),\ldots,g(v_n)\}\;\;\;\text{where}\;\;\;g(v_i) = h(\lceil v_i\cdot\delta\rceil)$$

\noindent The function $h$ is the hash function that is applied to the rounded result. After applying $f(V)$, each secret in the resulting set is proven to the server using the zero knowledge protocol. It is difficult to produce every single keystroke within the $\delta$ tolerance, therefore only a $\gamma$ fraction of the secrets must be successfully proven to the server.

\section{Results}
\label{sec:results}
The performance of the system is evaluated on two main metrics: precision (how easily a person can repeat his own typing pattern), and protection (how easily an imposter can forge the typing pattern). The performance of the application is tested with four participants, all with varying typing speeds, preferred hand layouts, and dominant hands. The weighted difference in average typing speed at which the target user was typing at is recorded as $\epsilon'$ and compared to the threshold level of $\epsilon=40$ in the zero-knowledge case and $\epsilon=50$ for the basic protocol. The consistency that the target user could replicate that result is measured as a percentage. The three other participants them attempt to forge the typing signature, having not seen the target user's test trials.  Identical words were selected as the control for both protocols in order to measure the variance between the two success rates.

Table \ref{table:res-basic} contains the results from this experiment en the Basic Protocol implementation that is described in Section~\ref{sec:basic-prot}. Trials were conducted between 20 and 40 times with varying word lengths. The values of the timings were compared directly, allowing a total error of $\epsilon=50$. The value $\epsilon^{'}$ in Table~\ref{table:res-basic} indicates the average error value that the target user scored. Some trends arise from these results; as word length increases, typing signatures become more unique and therefore, adversary forgery success rate drops significantly. However, this also affects the target user's ability to recreate his typing signature consistently. There appears to be a tradeoff between security and convenience which is due to choosing a constant $\epsilon$. Longer keystroke vectors amortise the allowed error over a larger number of keystrokes, therefore each keystroke must be more exact.

\begin{table}[h!]
\centering
\begin{tabular}{||c c c c||} 
 \hline
 Length of Name & $\epsilon^{'}$ & Success by User (\%) & Success by Imposter (\%) \\ [0.5ex] 
 \hline\hline
 3 & 14 & 100 & 100 \\ 
 5 & 16 & 87 & 65 \\
 8 & 47 & 85 & 25 \\ 
 11 & 43 & 87 & 12 \\
 13 & 47 & 65 & 0 \\ [1ex] 
 \hline
\end{tabular}
\caption{Basic Protocol: Keystroke variation by length of name.}
\label{table:res-basic}
\end{table}

Table~\ref{table:res-zkp} documents repeated test trials with $\delta=0.05$. Trials were conducted between 20 and 40 times with varying word lengths. The trend of increasing successful authentications with increasing word length is apparent. This is most likely attributed to the fact that typing patterns are more distinct with longer pieces of text. Conversely, imposters are also more likely to successfully impersonate the target user with a longer text, since the scheme does not measure the accumulated difference in time. In other words, an imposter's speed could be very far from the expected timings on many of the strokes, but but close on others. In this case, he may still be authenticated because the protocol only cares \textit{how many} matches occured, not \textit{how much} the totals differ. With shorter words, the implementation is very unforgiving.  Since there are fewer keystrokes to average out the mistakes, timing is extremely sensitive.  In practice, this is less of a concern since virtually no one has a combined first name and last name of less than three characters. There exists a non-trivial difference between the success rates of target users and their impersonators. This affirms the fact that keystroke dynamics are variable enough and have sufficient entropy to be used as authentication tools. Barring anomalies with short phrases, the difference in patterns become apparent. In particular, the increase between a phrase of 11 characters and one of 13 show that longer strings are not much harder for a user to type, but decrease the success rate of an adversary.

\begin{table}[h!]
\centering
\begin{tabular}{||c c c c||} 
 \hline
 Length of Name & $\epsilon^{'}$ & Success by User (\%) & Success by Imposter (\%) \\ [0.5ex] 
 \hline\hline
 3 & 112 & 4 & 0 \\ 
 5 & 59 & 10 & 5 \\
 8 & 47 & 85 & 23 \\ 
 11 & 34 & 87 & 43 \\
 13 & 32 & 87 & 35 \\ [1ex] 
 \hline
\end{tabular}
\caption{Zero-Knowledge Protocol: Keystroke variation by length of name.}
\label{table:res-zkp}
\end{table}

@@

In fact, an ideal protocol would combine the information-theoretic security of the zero-knowledge protocol with the protection against forgers of the basic protocol. The false login rate of the current implementation of the zero knowledge protocol is much too high. One way to achieve better results in a zero knowledge setting may be to use the fuzzy extractor construction presented in Section~\ref{sec:bg-fuzzy}. The results of this extractor (89\% genuine accept rate and 1\% false accept rate~\cite{conf/csreaSAM/AlvarezEA09}) show promise in preserving more of the information about the keystroke timings than the rounding extractor that is implemented. The direct comparison results in Table~\ref{table:res-basic} show that achieving a relatively high genuine acceptance rate is possible while keeping the false accept rate low in a keystroke dynamics authentication scheme. This result is reaffirmed by the experiments of Ara\'ujo, Sucupira, Lizarraga, Ling, and Yabu-Uti~\cite{Araujo:2005:UAT:2197882.2199456}. Since the purpose of a fuzzy extractor is to separate data from noise, the actual information stored in a keystroke vector should be able to be extracted almost in its entirety~\cite{conf/csreaSAM/AlvarezEA09}. There is no reason to believe that the direct comparator can differentiate data from noise better than a well-designed fuzzy extractor. In fact, in 2013, Fuller, Meng, and Reyzin created a fuzzy extractor construction that has no entropy loss~\cite{cryptoeprint:2013:416}. That is, the amount of entropy in the secret is equal to the amount of entropy in the input template. This result may not be desirable in all cases as some of the noise needs to be filtered out, but it shows that information loss is not inherent in the zero knowledge protocol.

\section{Future Work}
\label{sec:future}
It is evident that the implementation, though viable as a proof of concept, is nowhere near robust enough to be used in practice. Currently, for longer names, the prototype has bounded probability of success; it correctly verifies the correct user a significant portion of time (barring human induced errors such as typing anomalies), but also verifies an imposter over $40\%$ of the time.  This can be fixed by adjusting $\delta$, the fault tolerance per keystroke. The idea is to find the optimal value of $\delta$ and $\epsilon$ such that the real users can be easily validated but an imposter can not be validated. The next logical step would be to adjust the implementation to use a different fuzzy extractor such as the one described in Section~\ref{sec:bg-fuzzy}~\cite{Araujo:2005:UAT:2197882.2199456}. If sufficient performance is not achieved, then a fuzzy extractor with no entropy loss~\cite{cryptoeprint:2013:416} could be used.

Aside from accuracy concerns, the protocol presented is very keyboard dependent. This means that this method of verification would be impossible on mobile devices, since people generally use their thumbs to type on smaller keyboards and the keystroke timing on these keyboards is likely to be very different. All of the trials used a QWERTY keyboard the layout of which is significantly different than that of the French AZERTY keyboard or Dvorak. Although not yet tested, the protocol presented would work just as well on different keyboard layouts assuming that users can type with equal consistency on these keyboards. A problem arises when the keyboard layout is switched. For example, if an American user travels to France, he may be unable to access his account due to the different keyboard layout.

The protocol should be tested on many more words of varying lexicographical distances. The demonstration used only five words which is subject to extreme bias, since characters in the words chosen may have been clustered in a way that changed the difficulty of typing that one particular word. The currently observed trends may differ when a larger sample size of words and participants is used.

Though the implementation is not yet accurate enough, it provides a solid proof of concept for the potential of keystroke authentication. The protocol could be improved by augmenting the keystroke vector with additional characteristics of a person's typing pattern such as the actuation force and stroke duration. Because of the nature of our zero-knowledge scheme, the users also do not have to worry about any security breaches in the communication channels or in the database. Just like a written signature, the typing signature will be unique and secure.
